### 2018-03-30 reports/thesis/chapters/5_inference.tex
Started on mode finding.

The method will be described here. Suppose the following test statistics were obtained $z_1,z_2,\dotdotdot,z_n$. First the histogram needs to be smoothed. \cite{efron2004large} recommended using splines but a kernel density estimate \citep{parzen1962on, friedman2001elements} will be used here because it has simpler mathematical properties. The kernel density estimate is
\begin{equation}
	\widehat{p}_Z(z)=
	\frac{1}{nh}
	\sum_{i=1}^n\phi\left(
		\dfrac{z_i-z}{h}
	\right)
\end{equation}
where $h$ is the bandwidth. The bandwidth was chosen such that
\begin{equation}
	h = 0.9n^{-1/5}\times\text{min}\left(s_z,\text{iqr}_z/1.34\right) + 0.16
\end{equation}
where $s_z$ and $\text{iqr}_z$ are the standard deviation and interquartile range. The parameters of the empirical null are estimated using the kernel density estimate. The resulting estimated density is shown in Figure \ref{fig:inference_sub_density_estimate}.

@@

>>>>

The method will be described here. Suppose the following test statistics were obtained $z_1,z_2,\dotdotdot,z_n$. First the histogram needs to be smoothed. \cite{efron2004large} recommended using splines but a kernel density estimate \citep{parzen1962on, friedman2001elements} will be used here because it has simpler mathematical properties. The kernel density estimate is
\begin{equation}
	\widehat{p}_Z(z)=
	\frac{1}{nh}
	\sum_{i=1}^n\phi\left(
		\dfrac{z_i-z}{h}
	\right)
	\label{eq:inference_kernel_density_estimate}
\end{equation}
where $h$ is the bandwidth. The bandwidth was chosen such that
\begin{equation}
	h = 0.9n^{-1/5}\times\text{min}\left(s_z,\text{iqr}_z/1.34\right) + 0.16
\end{equation}
where $s_z$ and $\text{iqr}_z$ are the standard deviation and interquartile range. The parameters of the empirical null are estimated using the kernel density estimate. The resulting estimated density is shown in Figure \ref{fig:inference_sub_density_estimate}.

@@

The mode can be found by solving for $\widehat{\mu}_0 = \argmax\widehat{p}_Z(z)$. This cannot be solved in closed form easily and should be done numerically. The  Newton-Raphson method can be used to solve $\hat{p}_Z'(z) = 0$ for $z$. Instead the log space will be used, solving
\begin{equation}
	\dfrac{
		\partial
	}
	{
		\partial z
	}
	\ln\widehat{p}_Z(z)
	= 0
\end{equation}
for $z$. This will at least find stationary points.

The Newton-Raphson method is an iterative algorithm and requires an initial value $z^{(0)}$. The iterative step is
\begin{equation}
	z^{(r+1)} =
	z^{(r)}
	-\dfrac{
		\left.
			\dfrac{
				\partial
			}
			{
				\partial z
			}
			\ln\widehat{p}_Z(z)
		\right|_{z = z^{(r)}}
	}
	{
		\left.
			\dfrac{
				\partial^2
			}
			{
				\partial z^2
			}
			\ln\widehat{p}_Z(z)
		\right|_{z = z^{(r)}}
	} 
\end{equation}
for $r=0,1,2,3,\dotdotdot$ until some convergence condition is met. The derivatives for the log density can be obtained via the following. Recall that the kernel density estimate is
\begin{equation}
	\widehat{p}_Z(z)=
	\frac{1}{nh}
	\sum_{i=1}^n\phi\left(
		\dfrac{z_i-z}{h}
	\right)
	\tag{\ref{eq:inference_kernel_density_estimate}}
\end{equation}
so that the log density is
\begin{equation}
	\ln\widehat{p}_Z(z)=
	\ln\left(
		\dfrac{1}{nh}
	\right)
	+
	\ln\left[
		\sum_{i=1}^n
		\phi\left(
			\dfrac{
				z_i - z
			}
			{
				h
			}
		\right)
	\right]
	\ .
\end{equation}
Taking the first derivative
\begin{equation}
	\dfrac{
		\partial
	}
	{
		\partial z
	}
	\ln\widehat{p}_Z(z)
	=
	\dfrac{
		1
	}
	{
		\sum_{i=1}^n
		\phi\left(
			\dfrac{
				z_i - z
			}
			{
				h
			}
		\right)
	}
	\times
	\sum_{i=1}^n
	\phi'\left(
		\dfrac{
			z_i - z
		}
		{
			h
		}
	\right)
	\left(
		-\dfrac{
			1
		}
		{
			h
		}
	\right)
	\ .
\end{equation}
Using the fact that $\phi(z)=(2\pi)^{-1/2}\exp(-z^2/2)$, then $\phi'(z)=-z\phi(z)$. This can be used to simplify the equation to be
\begin{equation}
	\dfrac{
		\partial
	}
	{
		\partial z
	}
	\ln\widehat{p}_Z(z)
	=
	\dfrac{
		\sum_{i=1}^n
		\phi\left(
			\dfrac{
				z_i - z
			}
			{
				h
			}
		\right)
		\left(
			\dfrac{
				z_i - z
			}
			{
				h
			}
		\right)
	}
	{
		h
		\sum_{i=1}^n
		\phi\left(
			\dfrac{
				z_i - z
			}
			{
				h
			}
		\right)
	}
	\ .
\end{equation}

Taking the derivative again
\begin{multline}
	\dfrac{
		\partial^2
	}
	{
		\partial z^2
	}
	\ln\widehat{p}_Z(z)
	=
	\left[
		h\sum_{i=1}^n
		\phi\left(
			\dfrac{
				z_i-z
			}
			{
				h
			}
		\right)
	\right]^{-2}
	\times
	\left\{
		h\left[
			\sum_{i=1}^n
			\phi\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
		\right]
	\right.
	\\
	\left.
		\times
		\sum_{i=1}^n\left[
			\phi'\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
			\left(
				-\dfrac{
					1
				}
				{
					h
				}
			\right)
			\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
			+\phi\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
			\left(
				-\dfrac{
					1
				}
				{
					h
				}
			\right)
		\right]
	\right.
	\\
	\left.
		-
		\left[
			\sum_{i=1}^n
			\phi\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
			\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
		\right]
		\left[
			h\sum_{i=1}^n
			\phi'\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
			\left(
				-\dfrac{
					1
				}
				{
					h
				}
			\right)
		\right]
	\right\}
	\ .
\end{multline}
Using the fact that $\phi'(z)=-z\phi(z)$, then it can be simplified to
\begin{multline}
	\dfrac{
		\partial^2
	}
	{
		\partial z^2
	}
	\ln\widehat{p}_Z(z)
	=
	\left[
		h\sum_{i=1}^n
		\phi\left(
			\dfrac{
				z_i-z
			}
			{
				h
			}
		\right)
	\right]^{-2}
	\times
	\left\{
		\left[
			\sum_{i=1}^n
			\phi\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
		\right]
	\right.
	\\
	\left.
		\times
		\left[
			\sum_{i=1}^n
			\phi\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
			\left(
				\left(
					\dfrac{
						z_i-z
					}
					{
						h
					}
				\right)^2
				-1
			\right)
		\right]
		-
		\left[
			\sum_{i=1}^n
			\phi\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
		\right.
	\right.
	\\
	\left.
		\left.
			\left(
				\dfrac{
					z_i-z
				}
				{
					h
				}
			\right)
		\right]^2
	\right\}
	\ .
\end{multline}
